---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Set options (figure folder)
```{r}
library(knitr)
opts_chunk$set(fig.path = "figure/")
```

## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv", stringsAsFactors = FALSE)
```

## What is mean total number of steps taken per day?
Count total number of steps per day
```{r}
total_steps_per_day <- tapply(data$steps, data$date, sum, na.rm = TRUE)
```

And create a histogram for total number of steps per day
```{r}
hist(total_steps_per_day, xlab = "Total Steps per Day", main = "Histogram of Total Steps per Day")
```

Useful inforamtion about total steps per day
```{r}
summary(total_steps_per_day)
```

## What is the average daily activity pattern?
Average steps by interval
```{r}
average_steps_by_interval <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
```

And build a line chart by time interval with red line which indicates max value
```{r}
plot(names(average_steps_by_interval), average_steps_by_interval, type="l", xlab = "Time Interval", ylab = "Steps", main = "Average Steps by Daily Time Interval")
abline(v = names(which.max(average_steps_by_interval)), col = "red")
```

Interval with max value corresponds to next time
```{r}
max_interval = which.max(average_steps_by_interval)
max_interval
sub(pattern = "([0-9]{2})([0-9]{2})",
    replacement = "\\1:\\2",
    x = sprintf("%04s", names(max_interval)))
```

## Imputing missing values
```{r}
summary(data)
```
There are `r summary(data)[7]` missing values in steps.

I will use 5-minutes mean to fill in missed values.

To replace missing values with means - create array of mean values which corresponds to original array and use **is.na()** function as index to replace missing values only
```{r}
mean_data <- data
mean_steps <- rep_len(average_steps_by_interval, length(data$steps))
mean_data$steps[is.na(data$steps)] <- mean_steps[is.na(data$steps)]
```

There are head-s
```{r}
head(data)
head(mean_data)
head(average_steps_by_interval)
```

 and tail-s for used datasets
```{r}
tail(data)
tail(mean_data)
tail(average_steps_by_interval)
```

Count total number of mean steps per day
```{r}
total_mean_steps_per_day <- tapply(mean_data$steps, mean_data$date, sum, na.rm = TRUE)
```

And create a histogram for total number of mean steps per day
```{r}
hist(total_mean_steps_per_day, xlab = "Total Steps per Day", main = "Histogram of Total Mean Steps per Day")
```

Useful inforamtion about total mean steps per day
```{r}
summary(total_mean_steps_per_day)
```

Histogramm looks similary to the original one, except the first bucket which is much smaller.
3rd quantille in new data is the same, median is almost the same as in original data.weekda

## Are there differences in activity patterns between weekdays and weekends?

Mutate date column from character to Date and add a new factor column which identified weekdays or weekends
```{r}
mean_data$date <- as.Date(mean_data$date)
mean_data$weekday <- factor(rep_len("weekday", length(mean_data$date)), levels = c("weekday", "weekend"));
weekdays <- weekdays(mean_data$date);
mean_data$weekday[weekdays == "Sunday" | weekdays == "Saturday"] <- "weekend"
```

Lets validate results
```{r}
table(mean_data$weekday)
table(weekdays(mean_data$date))
```

And finally build a plot for daily activity for weekdays and weekends, I will use ggplot2 as graphics system and dplyr for groupping
```{r}
library(dplyr)
library(ggplot2)

by_interval_weekday <- group_by(mean_data, interval, weekday) %>%
                       summarize(steps = mean(steps))

qplot(x = interval, y = steps, geom = "line", data = by_interval_weekday, facets = weekday ~ . )
```
